---
title: "Problems_from_ch9"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 9.1

### Problem 9.1.1
```{r message=FALSE, warning=FALSE}
library(alr4)
```
```{r}
summary(Rpdata)
pairs(~. -y, Rpdata)
```
There is nothing strange.

### Problem 9.1.2
```{r}
# Fit the model
m1<-lm(y~.,Rpdata)
summary(m1)
```
There is nothing strange.

### Problem 9.1.3
```{r}
residualPlot(lm(y ~ ., data = Rpdata))
```

## Problem 9.11
```{r}
# Apply transformations and fit the model
fuel2001<-transform(fuel2001, Dlic = 1000 * Drivers/Pop,
                    Fuel = 1000 * FuelC/Pop,
                    Income = Income/1000)

m1 <- lm(Fuel~Tax+Dlic+Income+log(Miles), fuel2001)
# Create indexes
rows <- c("AK", "NY", "HI", "WY", "DC")
# No. of observations
n<-nrow(fuel2001)

# Get summary table
summ <- data.frame(cook = cooks.distance(m1)[rows],
                   Fuel = fuel2001[rows,"Fuel"],
                   residu = resid(m1)[rows],
                   h = hatvalues(m1)[rows])
# calculate standardized residual
summ$r <- summ$residu / (64.891*sqrt(1-summ$h))
# calculatet statistics for outlier test
summ$t <- summ$r*((n-5-1)/(n-5-summ$r^2))^(1/2)
# calculate Bonferroni corrected p???value
summ$p <- 5*2*pt(-abs(summ$t),46)
# Use 5 because we have to report for these 5 cases
summ
```

None of them would be declared outliers. 
Alaska has the largest influence on regression.

## Problem 9.15

### Problem 9.15.1
```{r}
m1 <- lm(Life ~ poly(Speed, 2) + poly(Feed, 2) + Speed:Feed, lathe1)
boxCox(m1)
```
This is the graph of the profile log-likelihood for the transformation parameter using the Box-Cox method for the second-order lathe model. The confidence interval for lambda???? is very narrow and includes zero, suggesting a log transformation.

### Problem 9.15.2
```{r}
m2 <- update(m1, log(Life) ~ .)
influenceIndexPlot(m2)
```
Observation 9-12 have large leverages by design in this planned experiment. These were "star points", unreplicated observations at relatively extreme values of the predictors to model curvature. Two of these, 9 and 10, also have large residuals and these in combination give large values for Cook's Distance as well.
```{r}
scatterplot(Speed ~ Feed, lathe1, id.method=9:12, smooth=FALSE, reg.line=FALSE,
boxplots=FALSE, cex=2, pch=20)
```
With all the data, the analysis of variance table is
```{r}
Anova(m2)
```
With the two cases deleted, the interaction is no longer important.
```{r}
m3 <- update(m2, subset=-c(9, 10))
Anova(m3)
```
In this example deleting the points is probably not called for; after all, this were designed to be influential cases, so there is a hint of an interaction here.

## Problem 9.16
Plot Buchanan vs Bush and label 'Palm Beach'. It can be seen that Palm Beach has an unusually high number of Buchanan votes. 
```{r}
aa <- florida[c("PALM BEACH"),]
with(florida, plot(Buchanan ~ Bush))
text(aa[,2:3], labels = row.names(aa), pos = 1)
```
Conduct an outlier test for the mean function E(Bucahanan | Bush)
```{r}
outlierTest(lm(Buchanan ~ Bush, florida), cutoff = 67, n.max = 67)
```
The outlier test suggests that Palm Beach is indeed an outlier.

From the plot we also see another country that has an unusual number of votes. The point on the right has an unusually high number of Bush votes.

To find the point:
```{r}
rownames(florida[florida$Bush > 250000,])
```

We see that the point is DADE county. We plot our data again, labeling our points.
```{r}
aa <- florida[c("PALM BEACH","DADE"),]
with(florida, plot(Buchanan ~ Bush))
text(aa[,2:3], labels = row.names(aa), pos = 1)
```

Transforming our data to log scale and performing Outlier Test:
```{r}
plot(log(florida$Buchanan) ~ log(florida$Bush), xlab="Log(Bush)", ylab="Log(Buchanan)")
```

```{r}
m <- lm(log(Buchanan)~log(Bush),florida)
outlierTest(m, cutoff=67,n.max=67)
```

From this table we can see that the value of DADE county does not appear to be an outlier on the logarithmic scale.