---
title: "Problems from chapter_2"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 2.4

### 2.4.1
Plotting y=bigmac2009 versus x=bigmac2003 with the ols fitted line and 
the y=x line.
```{r message=FALSE}
library(alr4)
plot(bigmac2009 ~ bigmac2003, data = UBSprices, col="lightblue", pch=19, cex=0.9, font=2)
text(bigmac2009 ~ bigmac2003, data = UBSprices, cex=0.9, font=2)
reg <- lm(bigmac2009 ~ bigmac2003, data = UBSprices)
abline(reg, col="red")
abline(0,1)

```
Unusual cases:
For most of the data points, the price of big mac in 2003 was more than the price in 2009. However, for points that are lying above the y=x line, the price in 2009 is greater than the price in 2003. Hence, the cases like points 45, 20, 11, 8, 31, 12, 21 etc are the unusual cases. 

### 2.4.2
We can see from the plot the data is neither showing linearity nor is it correlated. Also, the variance is varying a lot as we move across the x-axis.

### 2.4.3
Plotting log(bigmac2009) versus log(bigmac2003).
```{r}
plot(bigmac2009 ~ bigmac2003, log = "xy", data = UBSprices)
```
We can see that the data here is showing strong linearity alog with a strong correlation between x and y. As we move across the x-axis, the values of y are also increasing. The variance across x-axis also remains the same more or less, specially in the central portion of the data. Hence, is graph can be more sensibly summarized with a linear regression.

## Problem 2.5
```{r message=FALSE}
reg <- lm(Late ~ Early, data = ftcollinssnow)
summary(reg)
```
We can verify that the t-value fot the slope is 1.553.

## Problem 2.15

### 2.15.1
```{r message=FALSE}
library(alr4)
fit <- lm(Length ~ Age, data = wblake)
# 95% confidence intervals for mean length at ages 2, 4 and 6
p246 <- predict(fit, data.frame(Age=c(2,4,6)), interval = "confidence")
p246
```

### 2.15.2
```{r}
# 95% confidence intervals for mean length at age 9
p9 <- predict(fit, data.frame(Age=9), interval = "confidence")
p9
```
This interval is likely to be untrustworthy as the data does not have any fish with age greater than 8. So, age = 9 is an extrapolation of the data and we are ensure if our mean function is valid for age = 9 or not.

## Problem 2.16

### 2.16.1
```{r}
logfertility <- log(UN11$fertility)
logppgdp <- log(UN11$ppgdp)
reg <- lm(logfertility ~ logppgdp)
summary(reg)
```

### 2.16.2
```{r}
plot(logfertility ~ logppgdp, data = UN11)
abline(reg, col = "red")
```

### 2.16.3
The t-test can be used, with t = -14.79 with 197 degrees of freedom. Th p-value can be seen to be extremely small, so the one-sided p-value will also be close to zero. Thus, we have strong evidence that Beta1 < 0, suggesting that mostly countries with higher log(ppgdp) have lower log(fertility).

### 2.16.4
From the summary of the data we can see that the value of R-squared is 0.526, which is the coefficient of determination. It means that 52.6% of the variablity in log(Fertility) can be explained by conditioning log(ppgdp). 

### 2.16.5
```{r}
# 95% prediction interval for data with ppgdp = 1000 
p1000 <- predict(reg, data.frame(logppgdp=log(1000)), interval = "prediction")
p1000

# The point prediction is 1.234567

# Finding prediction interval for fertility by taking 
#exponential values of lower and upper bounds of p1000:
exp(0.6258791)
exp(1.843256)
```

### 2.16.6

```{r}
#Locality with highest value of fertility
rownames(UN11)[which.max(UN11$fertility)]

#Locality with lowest value of fertility
rownames(UN11)[which.min(UN11$fertility)]

#Two localities with largest positive residuals
xx <- sort(reg$residuals, decreasing = TRUE)
xx[1:2] # Gives indices 58 and 4
rownames(UN11)[58]
rownames(UN11)[4]

#Two localities with largest negative residuals
xy <- sort(reg$residuals)
xy[1:2] # Gives indices 23 and 118
rownames(UN11)[23]
rownames(UN11)[118]
```

## Problem 2.20

### 2.20.1

```{r}
reg1 <- lm(Interval ~ Duration, data = oldfaith)
summary(reg1)

```
Equation for predicting Interval from Duration:
Interval = 33.987808 + 0.176863(Duration)

### 2.20.2
```{r}
# 95% confidence intervals for duration 250 seconds
p250 <- predict(reg1, data.frame(Duration=250), interval = "prediction")
p250
```
We are 95% confident that the individual will have to wait between 66.35 and 90.05 seconds for the next eruption.

### 2.20.3
```{r}
p <- predict(reg1, data.frame(Duration=250))
p
# We find the mean to be 78.20354
# We find the standard deviation to be 6.004 from the summary. 
#Since we assume constant variance, we assume that the standard deviation is same throughout.

qnorm(0.9, mean = 78.20354, sd = 6.004)
```
Hence, 85.89798 is the 0.90 quantile.


