---
title: "Problems_from_ch6_and_ch7"
output: pdf_document
---

## Problem 6.3
```{r}
library(alr4)
mod1 <- lm(lifeExpF ~ group, data = UN11)
mod2 <- update(mod1, ~. + log(ppgdp))

anova(mod1, mod2)
```

We get the F-test for NH: model 6.7 vs AH: model 6.9. Since the p-value is very small, we conclude that we should reject the NH model.

## Problem 6.4

### Problem 6.4.1
The NH model includes log(ppgdp) as the regressor along with the interaction term for the interaction between group and log(ppgdp).

### Problem 6.4.2
```{r}
mod4 <- lm(lifeExpF ~ group + log(ppgdp) + group:log(ppgdp), data = UN11)
mod3 <- update(mod4, ~. - group)

anova(mod3, mod4)
```

Fail to reject NH from p-value.

## Problem 6.6

## Problem 6.7

### Problem 6.7.1
```{r}
# Transformations
fuel2001$Fuel <- with(fuel2001, 1000*FuelC/Pop)
fuel2001$Dlic <- with(fuel2001, 1000*Drivers/Pop)
fuel2001$Income <- with(fuel2001, Income/1000)

m1 <- lm(FuelC ~ Tax + Dlic + Income + log(Miles), data = fuel2001)
m2 <- lm(FuelC ~ log(Miles) + Income + Dlic + Tax, data = fuel2001)

anova(m1)
anova(m2)
```

### Problem 6.7.2
```{r}
Anova(m1)
Anova(m2)
```

## Problem 6.9
```{r}
w1 <- lm(Y ~ X1 + I(X1^2) + X2 + I(X2^2) + X1:X2, data = cakes)
w2 <- update(w1, ~. - X1:X2)
w3 <- update(w1, ~. - I(X1^2))
w4 <- update(w1, ~. - X1 - I(X1^2) - X1:X2)

anova(w2,w1)
anova(w3,w1)
anova(w4,w1)
```

In all three cases, p-values are smaller than 0.01. Hence, we can say that we reject the null hypothesis in all three cases.

## Problem 7.2

### Problem 7.8.1
```{r}
plot(Weight ~ Age, data = jevons)
```


```{r}
plot(SD ~ Age, data = jevons)
```

We can see that linearity holds but variability is not constant. Older coins seem more variable.

### Problem 7.8.2
```{r}
mm <- lm(Weight ~ Age, data = jevons, weights = n/SD^2)
summary(mm)
```

### Problem 7.8.3
```{r}
predict(mm, data.frame(Age=0), interval = "confidence")
```

